Progress with the NOSAMS<br>Hybrid Gas Ion Source

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Reorder first slides. Minor edits.

---
title: "Progress with the NOSAMS"
subtitle: "Hybrid Gas Ion Source"
author: "Brett Longworth"
institute: "NOSAMS/WHOI"
date: "2021/05/03"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

- Ultimate precision is lower
- Lower currents/counts (time to precision)
- variability in blank (current dependent)
- variability in sample performance
]

Performance = currents counts stability efficiency time

@@

<br>
.pull-left[
* Inject CO<sub>2</sub> gas onto Ti target
* Sputter with Cs<sup>+</sup>
* Dissociate molecules
* Produce C<sup>-</sup> ions (and O<sup>-</sup>, CH<sup>-</sup>, OH<sup>-</sup>, CO<sup>-</sup>...)
* Extract C<sup>-</sup> beam
]

@@

@@

Figure/photo of NEC bellows system

@@

.pull-left[
Potential for:

* Lower cost
* Better performance
]

@@

@@

@@

* increase partial pressure of CO<sub>2</sub> by dilution?
Gas inlet controls flow of CO<sub>2</sub> to source by varying pressure, volume, and gas composition.

Dilution with helium

Makes moving small amounts of gas easier by increasing pressure

@@

Introduce steady, reproducible current as key to data quality
Constant flow of CO<sub>2</sub> to source produces constant currents, which reduces ratio variability.

@@

### Flow control
* Change volume
* Add CO<sub>2</sub>
* Add Helium

@@

<br>
Smallest samples require minimizing inlet volume and pressure while maintaining good flow control
]

@@

# NOSAMS Closed inlet

@@

* Note similarity to first tests
* Flat top means stable currents/ratios for long enough to measure
* Sample size can be reduced to 2 mL (~1 mg C)
* Further reduction with increased flow to source

### Balance flow rates to maximize stable currents and minimize sample size
]

.pull-right[
<br>
<br>
<br>
```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Vial dilution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

@@

Modeled blank:
* Fm: `r format(blank_val$Fm_blank, digits = 3)`
* Current: `r format(blank_val$m_blank, digits = 3)` μA

Measured blank:
* Fm: `r ds %>% filter(Gas == "Blank") %>% pull(mean) %>% mean() %>% format(digits = 3)`
* Current: `r ds %>% filter(Gas == "Blank") %>% pull(Cur) %>% mean() %>% format(digits = 3)` μA

@@

### Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`

@@

@@

secs <- ggplot(carb_data, aes(fm_consensus, fm_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr), 
                  position = position_dodge2(width = dw),
                  size = ps,) +
  lims(x = c(-0.05, 1.1),
       y = c(-yl, yl)) +
  labs(title = "Difference from expected Fm",
       subtitle = "Uncorrected",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.direction = "horizontal",
        legend.position = c(.86, 1),
        legend.justification = "bottom")

secs_lbc <- ggplot(carb_data, aes(fm_consensus, fmbc_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr), 
                  position = position_dodge2(width = dw),
                  size = ps) +
  lims(x = c(-0.05, 1.1),
       y = c(-yl, yl)) +
  labs(subtitle = "Blank corrected",
       x = "Fm expected",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(legend.position = "None")

@@

#### Source differences
* one target in source at a time - lower blank
* gas supplied via conical joint in side of target

@@

Current goals:
* Finish characterization of the method
* Compare to graphite with test samples
* Improve accuracy
* Better understand method blank
* Reduce sample size
* Automation of data acquisition and reduction

>>>>

---
title: "Progress with the NOSAMS<br>Hybrid Gas Ion Source"
author: "Brett Longworth"
institute: "NOSAMS/WHOI"
date: "2021/05/03"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

@@

.pull-left[
* Inject CO<sub>2</sub> gas onto Ti target
* Sputter with Cs<sup>+</sup>
* Dissociate molecules
* Produce C<sup>-</sup> ions (and O<sup>-</sup>, CH<sup>-</sup>, OH<sup>-</sup>, CO<sup>-</sup>...)
* Extract C<sup>-</sup> beam
]

@@

Plasma in target recess? Blue glow.

- Ultimate precision is lower
- Lower currents/counts (time to precision)
- Variability in blank (current dependent)
- Variability in sample performance
]

@@

@@

.pull-left[
### Goal:

* Maximize gas density and residence time where sputtering and ionization are happening.

### Potential for:

* Better performance
* Lower cost
]

@@

CAMS and ETH have both done a lot of target geometry research.

In the end, compromise between simplicity and function.

Maybe still gains to be made. 
---

@@

Bottle gas allowed testing full range of flows

Precision not accuracy: modern gas normalized to itself

@@

Gas inlet controls flow of CO<sub>2</sub> to source by varying pressure, volume, and gas composition.

increase partial pressure of CO<sub>2</sub> by dilution?

Dilution Makes moving small amounts of gas easier by increasing pressure

@@

Introduce steady, reproducible current as key to data quality

Constant flow of CO<sub>2</sub> to source produces constant currents, which reduces ratio variability.

@@

### Flow control
* Change volume
* Add CO<sub>2</sub>
* Add helium

@@

@@

Smallest samples require minimizing inlet volume and pressure while maintaining good flow control
---

# NOSAMS closed inlet

@@

* Note similarity to first tests
* Flat top = stable currents/ratios 
* Sample size can be reduced to 2 mL (~1 mg C)
* Further reduction with increased flow to source

#### Balance flow rates to maximize stable currents and minimize sample size
]

.pull-right[
<br>
<br>
<br>
<br>
```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Vial dilution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

@@

.pull-left[
#### Modeled blank:
* Fm: `r format(blank_val$Fm_blank, digits = 3)`
* Current: `r format(blank_val$m_blank, digits = 3)` μA
]

.pull-right[
#### Measured blank:
* Fm: `r ds %>% filter(Gas == "Blank") %>% pull(mean) %>% mean() %>% format(digits = 3)`
* Current: `r ds %>% filter(Gas == "Blank") %>% pull(Cur) %>% mean() %>% format(digits = 3)` μA
]
]

@@

#### Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`

@@

We'll see what fit looks like on next slide.

@@

secs <- ggplot(carb_data, aes(fm_consensus, fm_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr), 
                  position = position_dodge2(width = dw),
                  size = ps,) +
  lims(x = c(-0.05, 1.1),
       y = c(-yl, yl)) +
  labs(title = "Difference from expected Fm",
       subtitle = "Uncorrected",
       colour = "Cur (μA)",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 18,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.direction = "horizontal",
        legend.position = c(.84, 1),
        legend.justification = "bottom")

secs_lbc <- ggplot(carb_data, aes(fm_consensus, fmbc_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr), 
                  position = position_dodge2(width = dw),
                  size = ps) +
  lims(x = c(-0.05, 1.1),
       y = c(-yl, yl)) +
  labs(subtitle = "Blank corrected",
       x = "Fm expected",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 18,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(legend.position = "None")

@@

#### Source differences
* One target in source at a time - lower blank
* Gas supplied via conical joint in side of target

@@

###Goals:
* Finish characterization of the method
* Compare to graphite with test samples
* Improve accuracy
* Better understand method blank
* Reduce sample size
* Automate data acquisition and reduction
